### 2020-01-02 Pseudo Code/template.tex
Create template.tex

%!TEX program = Xelatex
\documentclass{article}
\usepackage{amsmath,amscd,amsbsy,amssymb,latexsym,url,bm,amsthm}
\usepackage{epsfig,graphicx,subfigure}
\usepackage{enumitem,balance,mathtools}
\usepackage{wrapfig}
\usepackage{mathrsfs, euscript}
\usepackage{hyperref}
\usepackage{listings}
\usepackage[ruled,lined,boxed,linesnumbered]{algorithm2e}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{exercise}{Exercise}[section]
\newtheorem*{solution}{Solution}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\postscript}[2]
    {\setlength{\epsfxsize}{#2\hsize}
    \centerline{\epsfbox{#1}}}
\renewcommand{\baselinestretch}{1.0}

\setlength{\oddsidemargin}{-0.365in}
\setlength{\evensidemargin}{-0.365in}
\setlength{\topmargin}{-0.3in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\textheight}{10.1in}
\setlength{\textwidth}{7in}

\begin{document}

% ------------------ your name and number ----------------- %

% --------------------------------------------------------- %
\begin{itemize}
\item $U$: used classifiers $(d_{i})$
\item $R_i$: coding results: $(r_{i_1},r_{i_2},\cdots,r_{i_{n}}),r_i\in\{0,1\}$
\item $Pos_i$: Remaining labels if result of $d_i$ is $1$
\item $Neg_i$: Remaining labels if result of $d_i$ is $0$
\item $C$: array of coding results $(R_1,R_2,\cdots,R_m)$
\item $E$: Remaining labels of each coding in $C$: $(\{l_j\}),|C|=|R|$
\item $m_i$: possibility of result $r_i$:
$$
m_i=\left\{
\begin{array}{cc}
\frac{2}{K} & r_i=1, positive\\
\frac{K-2}{K} & r_i=0, negative\\
1 & r_i=2, meaningless\\
\end{array}
\right.
$$

\item $a=(a_1,a_2)$: concatenation of two arrays

\end{itemize}

\begin{algorithm}
\caption{}
$U=\phi$\\
$Entropy\leftarrow \sum_{i=1}^{L}\frac{1}{L}\log\frac{1}{L}=L$\\
\While {$Entropy>0$}{
	$min\leftarrow\infty$\\
	\ForEach {$d_i\in D-U$}{
			$S,T\leftarrow Split(E,C,d_i)$\\

			\ForEach{$i\in\{1,2,\cdots,|S|\}$}{
				$p_i=\prod_{j=1}^{|T_i|}m_{T_i^j}$\\
			}
			$e_i\leftarrow\sum_{j=1}^{|S_i|}\frac{1}{|S_i|}\log\frac{1}{|S_i|}=S_i$\\
			$H\leftarrow\sum_{j=1}^{|S|}p_je_j$\\
			\If{$H<min$}{
				$S_{min}\leftarrow S$\\
				$T_{min}\leftarrow T$\\
				$k\leftarrow i$\\
				$min\leftarrow H$\\
			}
	}
	$Entropy\leftarrow min$\\
	$E\leftarrow S_{min}$\\
	$C\leftarrow T_{min}$\\
	$U\leftarrow (U,d_k)$\\
}

\end{algorithm}

\begin{algorithm}
\caption{$Split(E,C,d_i)$}
$S\leftarrow$ Empty Array\\
$T\leftarrow$ Empty Array\\
\ForEach{$E_i\in E$}{
	\If {$Pos_i\cap E_i=\phi$ or $Neg_i\cap E_i=\phi$}{
		$S\leftarrow S+\{E_i\}$\\
		$T_0\leftarrow (C_i,2)$\\
		$T\leftarrow (T,(T_0))$\\
	}
	\Else{
		$S\leftarrow S+\{E_i\}\cap Pos_i$\\
		$T_0\leftarrow (C_i,1)$\\
		$T\leftarrow (T,(T_0))$\\
		$S\leftarrow S+\{E_i\}\cap Neg_i$\\
		$T_0\leftarrow (C_i,0)$\\
		$T\leftarrow (T,(T_0))$\\
	}
\Return $S,T$\\
}

\end{algorithm}

\end{document}