### 2020-01-03 Pseudo Code/template.tex
Update template.tex

%!TEX program = Xelatex
\documentclass{article}
\usepackage{amsmath,amscd,amsbsy,amssymb,latexsym,url,bm,amsthm}
\usepackage{epsfig,graphicx,subfigure}
\usepackage{enumitem,balance,mathtools}
\usepackage{wrapfig}
\usepackage{mathrsfs, euscript}
\usepackage{hyperref}
\usepackage{listings}
\usepackage[ruled,lined,boxed,linesnumbered]{algorithm2e}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{exercise}{Exercise}[section]
\newtheorem*{solution}{Solution}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\postscript}[2]
    {\setlength{\epsfxsize}{#2\hsize}
    \centerline{\epsfbox{#1}}}
\renewcommand{\baselinestretch}{1.0}

\setlength{\oddsidemargin}{-0.365in}
\setlength{\evensidemargin}{-0.365in}
\setlength{\topmargin}{-0.3in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\textheight}{10.1in}
\setlength{\textwidth}{7in}

\begin{document}

% ------------------ your name and number ----------------- %

% --------------------------------------------------------- %
\begin{itemize}
\item $U$: used classifiers: $(d_{i})$

\item $Pos_i$: Remaining labels if result of $d_i$ is $1$
\item $Neg_i$: Remaining labels if result of $d_i$ is $0$
\item $R_i$: code: array $(r_{i_1},r_{i_2},\cdots,r_{i_{n}}),r_i\in\{0,1,2\}$
$$
r_i=\left\{
\begin{array}{cc}
1 & result\ of\ d_i\ is\ true\\
0 & result\ of\ d_i\ is\ false\\
2 & d_i\ can\ only\ be\ true\ or\ false\ theoretically\\
\end{array}
\right.
$$
\item $C$: array of codes $(R_1,R_2,\cdots,R_m)$
\item $E$: Remaining labels of each code in $C$: $(\{l_j\}),|C|=|R|$

\end{itemize}

\begin{algorithm}
\caption{}
$U=\phi$;\\
$Entropy\leftarrow -\sum_{i=1}^{L}\frac{1}{L}\log\frac{1}{L}=logL$;\\
\While {$Entropy>0$}{
	$min\leftarrow\infty$;\\
	\ForEach {$d_i\in D-U$}{
			$S,T\leftarrow Split(E,C,d_i)$\tcp*[r]{if we choose $d_i$ as new classifier, renew the code and result}

			\ForEach{$i\in\{1,2,\cdots,|S|\}$}{
				$p_i=\prod_{j=1}^{|T_i|}m_{T_i^j}$ \tcp*[r]{probability of result $T_i$}
			}
			$e_i\leftarrow-\sum_{j=1}^{|S_i|}\frac{1}{|S_i|}\log\frac{1}{|S_i|}$;\\
			$H\leftarrow\sum_{j=1}^{|S|}p_je_j$;\\
			\If{$H<min$}{
				\tcc{record the min value of entropy and args}
				$S_{min}\leftarrow S$;\\
				$T_{min}\leftarrow T$;\\
				$k\leftarrow i$;\\
				$min\leftarrow H$;\\
			}
	}
	\tcc{renew the min value of entropy and args}
	$Entropy\leftarrow min$;\\
	$E\leftarrow S_{min}$;\\
	$C\leftarrow T_{min}$;\\
	add $d_k$ to used classifiers set $U$;\\
}

\end{algorithm}

\begin{algorithm}
\caption{$Split(E,C,d_i)$}
$S\leftarrow$ Empty Array;\\
$T\leftarrow$ Empty Array;\\
\ForEach{$E_i\in E$}{
	\If {$Pos_i\cap E_i=\phi$ or $Neg_i\cap E_i=\phi$}{
		$S\leftarrow S+\{E_i\}$\tcp*[r]{add $E_i$ to $S$ as an item directly}
		add a $2$ to the end of code $C_i$, and add the new code item to $T$;\\
	}
	\Else{

		$S\leftarrow S+(\{E_i\}\cap Pos_i)$\tcp*[r]{Split $E_i$ into two items, add the two items to $S$ separately}
		$S\leftarrow S+(\{E_i\}\cap Neg_i)$;\\
		add an $1$ to the end of code $C_i$, and add the new code item to $T$;\\
		add a $0$ to the end of code $C_i$, and add the new code item to $T$;\\

	}
\Return $S,T$;\\
}

\end{algorithm}

\begin{algorithm}
\caption{Heuristic-Search(U,C,E)}
\If {$|U|=k$}{
	\If {$|E_i|=1,\forall i\in\{1,2,\cdots,|E|\}$}
	{
		End the search and $U$ are the chosen classifiers\tcp*[r]{one code maps to one label}
	}
	\Else{
		\Return to the previous recursion;\\
	}
}
$PriorityQueue\leftarrow empty$\tcp*[r]{the priority queue(big-end) to record the entropy}
\ForEach {$d_i\in D-U$}{
	$E',C'\leftarrow Split(E,C,d_i)$\tcp*[r]{if choose $d_i$ as next classifier, renew codes and mapped labels}
	\ForEach{$j\in\{1,2,\cdots,|C|\}$}{
		$p_j=p(C_j)=p(r_1,r_2,\cdots,r_{|C_j|})$\tcp*[r]{probability of result $C_j'$}
	}
	$H'=\sum_{j=1}^{|E|}p_j(-\sum_{k=1}^{|E_j|}\frac{1}{|E_j|}\log\frac{1}{|E_j|})=\sum_{j=1}^{|E|}p_j\log |E_j|$;\\
	add $((E',C',k),H')$ to $PriorityQueue$ as $(key,value)$ pair;
}

\While{$PriorityQueue$ not empty}{
	$((E',C',k),H')\leftarrow$ top of $PriorityQueue$\tcp*[r]{fetch the element on the top of priority queue}
	Heuristic-Search($U\cup\{d_k\},C',E'$);
}

\end{algorithm}

\end{document}